The promise of community-driven preprints in ecology and evolution

### 2024-10-15 ms/ms.qmd
C11
C10

Gender differences in preprint use and publication outcomes have also been observed in several research fields, including ecology and evolutionary biology [@fox2019gender; @wehner2020comparison]. Therefore, such discrepancies are expected to manifest in preprint use on *EcoEvoRxiv*, but it is unclear to what extent. Understanding gender publishing patterns is challenging with observational data such as ours because we cannot know the gender of authors for certain, but we can use a data-driven approach to ascertain the probability that a particular name is of a given gender (man or woman). To obtain a rough idea of an author's gender, we used the R package *gender* (v.`r utils::packageVersion("gender")`; @mullen2021predict) to predict the most likely gender of the submitting author of a preprint. We used an algorithm to assign gender based on the submitting author’s name. We only used the algorithm-assigned gender when the gender of a given name was identified with 95% certainty. For the remaining names, we performed manual searches to determine gender based on the pronouns and photographs from professional and personal websites. We acknowledge that our approach does not capture self-assigned and non-binary genders. As such, our assumptions about an author's gender identity may be incorrect. Our data on gender had only two missing values--one where the first name of the submitting author was missing and the other one for a collective submission. As expected, we found that women were less likely to post to *EcoEvoRxiv* compared to men (women: `r (table(data4$gender)[1] / sum(table(data4$gender))*100)`%), reinforcing existing disparities between male and female scientists. For example, studies have shown that female first authors have lower acceptance rates and are cited less (~2%) compared to males [e.g., @fox2019gender].

@@

Despite the early successes of the new initiatives taken by *EcoEvoRxiv*, as described above, much work remains to be done to improve the understanding and use of pre- and postprints within our community. We view this perspective piece as a small step towards achieving that goal. We hope that readers are more familiar with the benefits of using community-driven preprint servers and the unique initiatives they can pursue. Community-driven preprint servers can set their own agenda and are driven by the needs and desires of the community. Supporting these initiatives should be a priority for all researchers. Volunteers at *EcoEvoRxiv* are encouraged to remain open to new and innovative ways to improve publication and open science practices. We believe that the future of preprints is bright, and community-driven initiatives, such as *EcoEvoRxiv* will play a crucial role in the future of scientific publishing.

>>>>

Gender differences in preprint use and publication outcomes have also been observed in several research fields, including ecology and evolutionary biology [@fox2019gender; @wehner2020comparison]. For example, gender gaps in preprint submissions were observed during COVID-19 lockdowns [@ucar2022mind], and previous surveys have shown that female participants are less likely to suggest posting articles as preprints, suggesting gender differences in views around preprints [@fraser2022motivations]. Therefore, such discrepancies are expected to manifest in preprint use on *EcoEvoRxiv*, but it is unclear to what extent. Understanding gender publishing patterns is challenging with observational data such as ours because we cannot know the gender of authors for certain, but we can use a data-driven approach to ascertain the probability that a particular name is of a given gender (man or woman). To obtain a rough idea of an author's gender, we used the R package *gender* (v.`r utils::packageVersion("gender")`; @mullen2021predict) to predict the most likely gender of the submitting author of a preprint. We used an algorithm to assign gender based on the submitting author’s name. We only used the algorithm-assigned gender when the gender of a given name was identified with 95% certainty. For the remaining names, we performed manual searches to determine gender based on the pronouns and photographs from professional and personal websites. We acknowledge that our approach does not capture self-assigned and non-binary genders. As such, our assumptions about an author's gender identity may be incorrect. Our data on gender had only two missing values--one where the first name of the submitting author was missing and the other one for a collective submission. As expected, we found that women were less likely to post to *EcoEvoRxiv* compared to men (women: `r (table(data4$gender)[1] / sum(table(data4$gender))*100)`%), reinforcing existing disparities between male and female scientists. For example, studies have shown that female first authors have lower acceptance rates and are cited less (~2%) compared to males [e.g., @fox2019gender].

@@

Despite the early successes of the new initiatives taken by *EcoEvoRxiv*, as described above, much work remains to be done to improve the understanding and use of pre- and postprints within our community. We view this perspective piece as a small step towards achieving that goal. We hope that readers are more familiar with the benefits of using community-driven preprint servers and the unique initiatives they can pursue. Community-driven preprint servers can set their own agenda and are driven by the needs and desires of the community. Supporting these initiatives should be a priority for all researchers. Volunteers at *EcoEvoRxiv* are encouraged to remain open to new and innovative ways to improve publication and open science practices. Our analysis can be used to drive changes in *EcoEvoRxiv* to make it a better platform for our community. We believe that the future of preprints is bright, and community-driven initiatives, such as *EcoEvoRxiv* will play a crucial role in the future of scientific publishing.
